---
title: "Homework 9"
author: "Lauren Maurer"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---
EID: lem4336

Link to GitHub: \url{https://github.com/laurenmaurer4/SDS_315.git}
```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE, results = 'hide')

library(dbplyr)
library(tidyverse)
library(mosaic)
library(MatchIt)
library(ggplot2)
library(kableExtra)

turnout <- read.csv("C:/Users/laure/Downloads/turnout.csv")
solder <- read.csv("C:/Users/laure/Downloads/solder.csv")
```



# Problem 1: Get Out the Vote

## Part A
```{r}
# The proportion of those receiving a GOTV call who voted in 1998.
got_call <- filter(turnout, GOTV_call == 1)
call_voted_1998 <- filter(got_call, voted1998 == 1)
call_prop_vote <- nrow(call_voted_1998)/ nrow(got_call)

# The sample proportion of those not receiving a GOTV call who voted in 1998
no_call <- filter(turnout, GOTV_call == 0)
no_call_voted_1998 <- filter(no_call, voted1998 == 1)
no_call_prop_vote <- nrow(no_call_voted_1998)/ nrow(no_call) 

# 95% confidence interval for the difference in these two proportions
confit <- prop.test(voted1998 ~ GOTV_call, data = turnout)
```

The proportion of those who received a GOTV call who voted in 1998 is `r round(call_prop_vote, 4)`. The sample proportion of those who did not receive a GOTV call and voted in 1998 in `r round(no_call_prop_vote, 4)`. When a 95% confidence interval is run for the difference in the two proportions the lower value is 0.1411 and the upper value is 0.2659.


## Part B
```{r}
con_MP_98<- xtabs(~ MAJORPTY + voted1998, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)

con_MP_call <- xtabs(~ MAJORPTY + GOTV_call, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)


con_96_98 <- xtabs(~ voted1996 + voted1998, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)
con_96_call <- xtabs(~ voted1996 + GOTV_call, data = turnout) %>%
  prop.table(margin = 2) %>%
  round(4)

```
### Variable: Major Party 
*0: False, 1: True*\

**Table 1: Proportions of Major Party and Voted in 1998** 
```{r, results='markup'}
con_MP_98
```

70% of people who did not vote in 1998 were part of a major party and 80% of people who voted in 1998 were part of a major party. Being part of a major party is a confounder because a higher proportion of the people in a major party voted in 1998.\

**Table 2: Proportions of Major Party and Got a GOTV Call**
```{r, results='markup'}
con_MP_call
```
74% of people who did not get a call were registered in a major party and 80% of the people who were called are also registered as a member in a major party. This is a confounder because there is a higher proportion of people who got the call and are part of a major party compared to people who did not get called and are part of a major party.\

### Variable: Voted in 1996 
\
*0: False, 1: True*\

**Table 3: Proportions of Voted in 1996 and Voted in 1998**
```{r, results='markup'}
con_96_98
```

Voting in 1996 is a confounder because people who voted in 1996 were more likely to vote in 1998.

**Table 4: Proportions of Voted in 1996 and Got a GOTV Call**
```{r, results='markup'}
con_96_call
```

Voting in 1996 is also related to people who got the GOTV call. The people who voted in 1996 were more likely to get the call.

### Variable: Major Party 
\
*0: Did not vote in 1998, 1: Voted in 1998*\

**Table 5: Mean Age of Voters in 1998**
```{r, results='markup'}
mean(AGE ~ voted1998, data = turnout) %>%
  round(4)
```
The people who voted in 1998 had a higher mean age than those who did not.

**Table 6: Mean Age of People Who Got GOTV Call**
```{r, results='markup'}
mean(AGE ~ GOTV_call, data = turnout) %>%
  round(4)
```

The mean age of people who got the call were older. The age of voters is a confounding variable because getting the call is related to age and so is voting in 1998.

## Part C
```{r}
# Matching data
matched_turnout <- matchit(GOTV_call ~ factor(voted1996) + AGE + factor(MAJORPTY), data = turnout, ratio = 5) %>%
  match.data
```

```{r, results='markup'}
# Proving there aren't confounders
xtabs(~ MAJORPTY + GOTV_call, data = matched_turnout) %>%
  prop.table(margin = 2) %>%
  round(4)

xtabs(~ voted1996 + GOTV_call, data = matched_turnout) %>%
  prop.table(margin = 2) %>%
  round(4)

mean(AGE ~ GOTV_call, data = matched_turnout)

```


All of the rows contain similar proportions indicating that the data is matched and has gotten rid of confounding variables

```{r}
# The proportion of those receiving a GOTV call who voted in 1998 for matched data.
got_call <- filter(matched_turnout, GOTV_call == 1)
call_voted_1998 <- filter(got_call, voted1998 == 1)
call_prop_vote_matched <- nrow(call_voted_1998)/ nrow(got_call)

# The sample proportion of those not receiving a GOTV call who voted in 1998 for matched data
no_call <- filter(matched_turnout, GOTV_call == 0)
no_call_voted_1998 <- filter(no_call, voted1998 == 1)
no_call_prop_vote_matched <- nrow(no_call_voted_1998)/ nrow(no_call) 

# 95% confidence interval for the difference in these two proportions
matched_confint <- prop.test(voted1998 ~ GOTV_call, data = matched_turnout)

```

For the matched data set, the proportion of people who received a GOTV call and voted is `r round(call_prop_vote_matched, 4)` and the sample proportion of people who did not receive a GOVT call and voted in 1998 is `r round(no_call_prop_vote_matched, 4)`The confidence interval is significant because it doesn't include zero so the conclusion is that when people receive a call they are more likely to vote. The lower value of the interval is 0.0056 and the upper value is 0.1418.


# Problem 2: Manufacturing Falws in Circuit Boards

## Part A
```{r}
opening_solder <- mutate(solder, Opening = factor(Opening))

ggplot(opening_solder) +
  geom_boxplot(aes(x=Opening, y=skips)) + labs(y='Skips', title = 'Opening Type vs. Number of Skips')
```
The graph above shows the opening type vs the range of number of skips. (L is large, M is medium, S is small)
```{r}
thickness_solder <- mutate(solder, Solder = factor(Solder))

ggplot(thickness_solder) +
  geom_boxplot(aes(x=Solder, y=skips)) + labs(y='Skips', title = 'Solder Thickness vs. Number of Skips')
```
The graph above shows the solder thickness vs the range of number of skips.

## Part B
```{r}
solder_model = lm(skips ~ Opening + Solder + Opening:Solder, data = solder)
solder_model_coef = coef(solder_model)

solder_model_coef_df <- data.frame(
  Coefficient_Value = solder_model_coef
)
row.names(solder_model_coef_df) <- c('Large Opening & Thick', 'Medium Opening', 'Small Opening', 'Thin', 'Medium Opening + Thin', 'Small Opening + Thin')



confint_solder_model = confint(solder_model, level = 0.95) %>% round(3)
row.names(confint_solder_model) <- c('Large Opening & Thick', 'Medium Opening', 'Small Opening', 'Thin', 'Medium Opening + Thin', 'Small Opening + Thin')
```
**Table of Estimate Coefficient**
`r kable(solder_model_coef_df)`\newpage

**Table of 95% Confidence Interval for Each Coefficient**
`r kable(confint_solder_model)`

## Part C

**Large Opening & Thick:** The baseline number of skips for the combination of large opening and thick solder is estimated to be 0.393\
**Medium Opening:** The main effect for the medium opening variable is 2.407. This is the effect of medium opening in isolation.\
**Small Opening:** The main effect for the small opening variable is 5.127. This is the effect of small opening in isolation.\
**Thin:** The main effect for the thin solder variable is 2.280. This is the effect of thin solder in isolation. \
**Medium Opening + Thin:** The interaction effect for medium opening and thin solder is -0.740.\
**Small Opening + Thin:** The interaction effect for small opening and thin solder is 9.653.\

## Part D

The opening size and thickness I would recommend to AT&T is medium opening and thin solder. The data shows that the combination of medium opening and thin solder will be 0.740 below the baseline number of skips which is 0.393.
