---
title: "Homework 10"
author: "Lauren Maurer"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

EID: lem4336

Link to GitHub: \url{https://github.com/laurenmaurer4/SDS_315.git}
```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE, results = 'hide')

library(dbplyr)
library(tidyverse)
library(mosaic)
library(MatchIt)
library(ggplot2)
library(kableExtra)
library(moderndive)
library(effectsize)

redline <- read.csv("C:/Users/laure/Downloads/redlining.csv")
groceries <- read.csv("C:/Users/laure/Downloads/groceries.csv")
```

# Problem 1: Redlining

```{r}
redline_lm = lm(policies ~ minority + fire + age + income, data = redline)
redline_table = get_regression_table(redline_lm)


standardize_parameters(redline_lm)
```

**Question:** The question that we are trying to answer is whether there is an association between the number of FAIR policies and that racial/ethnic composition of a ZIP code. 

**Approach:** A linear regression was used to answer the question. To account for confounding variables we adjusted for fire, age, and income.

**Results:** `r kable(redline_table, caption='Redlining Regression Table')`
```{r, message=FALSE}
ggplot(redline) + geom_histogram(aes(x=policies)) + labs(x='Number of Policies per 100 People',y='Count', title='Distribution of Number of Policies')
```
\newline
\newline
**Conclusion:** There is an association between the number of FAIR policies and the racial/ethnic composition of a ZIP code. When there is a 1% change in minority there is a 0.008 increase in the number of FAIR policies per 100 residents in a ZIP code. The 95% confidence interval for the change in policies is between 0.003 and 0.014. This isn't a significant change but also  not a small change. In the data, the number of policies per 100 people ranges from zero to a little over 2. If the minority population increases by 50% then there are 0.4 policies per 100 people. That increase covers a little under 1/5 of the range of all policies per 100 people in the data set.



# Problem 2: Grocery Store Prices

### Part A
```{r}
# Part A:
groceries_store_avg <- groceries %>%
  group_by(Store) %>%
  summarize(avg_price = mean(Price))

ggplot(groceries_store_avg) + geom_bar(aes(x=Store, y=avg_price), stat = 'identity') + labs(y='Average Prices of Products ($)', title = 'Store vs. Average Product Price', caption = str_wrap('The average price of all products in this plot range from 2 to 4. Whole Foods and Wheatsville Food Co-op have the highest average. Fiesta and Walmart have the lowest averages.')) + theme(plot.caption.position = "plot", plot.caption = element_text(hjust = 0)) + coord_flip()
```

### Part B
```{r, fig.height=6, fig.width=7}
# Part B

groceries_product <- groceries %>%
  group_by(Product) %>%
  summarise(num_stores = n())

ggplot(groceries_product) + geom_bar(aes(x=num_stores, y=Product), stat = 'identity') + theme(text = element_text(size = 12), plot.margin = margin(.005,.5,.005,.5, "cm")) + labs(x = "Number of Stores", title = 'Number of Stores a Product Is Sold In', caption = 'The plot above shows the number of stores that sell each product. Eggs and milk are sold in all 16 stores.') + theme(plot.caption.position = "plot", plot.caption = element_text(hjust = 0))
```

### Part C
```{r}
# Part C

groceries_lm = lm(Price ~ Product + Type, data = groceries)
groceries_lm_table = get_regression_table(groceries_lm)
```

Compared with ordinary grocery stores (like Albertsons, H-E-B, or Krogers), convenience stores charge somewhere between 0.41 and 0.92 dollars more for the same product.

### Part D
```{r}
# Part D

groceries_store_lm = lm(Price ~ Product + Store, data = groceries)
groceries_store_lm_table = get_regression_table(groceries_store_lm)
```

Walmart and Kroger Fresh Fare charge the least when comparing the same product. Whole Foods and Wheatsville Food Co-op charge the most when comparing the same product.

### Part E
```{r}
# Part E
```

H-E-B and Central Market are similar. Central Market charges an estimated \$0.07 more for the same product. This is a small difference compared to Central Market and Whole Foods where Whole Foods charges an estimated \$0.94 more for the same product.


### Part F
```{r}
# Part F

groceries <- groceries %>%
  mutate(Income10K = Income%/%10000)

groceries_income_lm = lm(Price ~ Product + Income10K, data = groceries)
groceries_income_lm_table = get_regression_table(groceries_income_lm)

standardize_parameters(groceries_income_lm)
```

As income increases by 10K the amount that people spend on food decreases by $0.13. People in poorer ZIP codes on average pay more for the same product. A one-standard deviation increase in the income of a ZIP code seems to be associated with a -0.03 standard-deviation change in the price that consumers in that ZIP code expect to pay for the same product.


